---
title: "Assignment 3c"
output: html_document
author: Animesh Sharma (as3592),Yaniv Bronshtein(yb262),Wanying Mo(wm318), Venkata Krishnam Raju Dalta(vkd20), Vipul Gharde(vig4), Aditya Maheshwari(am2971), Toshitt Ahuja(ta498), Fan Shen(fs470)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the ks.test, Durbin approach

Build a goodness of fit test for each distribution using the parametric bootstrap based on maximum likelihood

for every instance in the MOM and MLE estimator program

Test using a sample size of 100 for each distribution

Make sure to test against both the correct distribution and an one of the wrong distributions! for each case

```{r}
# library(graphics)
require(graphics)

```
**Let's look at an example of ks.test()**
```{r}
x <- rnorm(50)
y <- runif(30)

```


**Verdict: x and y do not come from the same distribution**
```{r}
# Do x and y come from the same distribution?
ks.test(x, y)
```
```{r}
# Does x come from a shifted gamma distribution with shape 3 and rate 2?
ks.test(x+2, "pgamma", 3, 2) # two-sided, exact
```

```{r}
ks.test(x+2, "pgamma", 3, 2, exact = FALSE)
```

H0: x+2 not greater than gamma distribution with shape 3 and rate 2
H1: x+2 is greater than gamma distribution with shape 3 and rate 2
Verdict: Fail to reject null hypothesis
```{r}
ks.test(x+2, "pgamma", 3, 2, alternative = "gr")
```

stochastic order quantifies the concept of one random variable being "bigger" than another.

```{r}
# test if x is stochastically larger than x2
x2 <- rnorm(50, -1) #mean=-1, sd=1
```

**Let's plot the ecdf() function for x. Empirical Distribution Function**
```{r}
par(mfrow=c(1,2))
plot(ecdf(x), xlim = range(c(x, x2)))
plot(ecdf(x2), ann = TRUE, lty = "dashed")

```

**Let's plot the ecdf() function for x2. Empirical Distribution Function**
```{r}
plot(ecdf(x2), ann = TRUE, lty = "dashed")

```

**T-test between x and x2**
**Verdict: Reject null hypothesis because mean(x) = -.06461931 and mean(x2)=-.99139101
```{r}
t.test(x, x2, alternative = "g")
```

**Wilcox Test between x and x2**
```{r}
wilcox.test(x, x2, alternative = "g")
```

```{r}
ks.test(x, x2, alternative = "l")
```
 
```{r}
ks.test(x, x2, alternative = "gr")
```