---
title: "Probability and Inference HW 3b"
author: "Yaniv Bronshtein(yb262), Vipul Ishwarchandra Gharde(vig4), Animesh Sharma(as3592),Wanying Mo(wm318), Fan Shen(FS470), Toshitt Ahuja(ta498), Venkata Datla(vkd 20), Aditya Maheshwari(am2971)"
date: "11/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Import the necessary libraries**
```{r}
library(purrr) 
library(MASS)
library(DIRECT)
```

**Functions to calculate first and second derivatives for Newton Raphson method**
```{r}
f1 <- function(a,n,x){
  return(-n*digamma(a)- n*log(mean(x)) + n*log(a) + sum(log(x)))
}
f2 <- function(a,n){
  return(-n*trigamma(a) + n/a)
}
```

**This function is used to calculate parameters for**
**different distributions of data using mean and variance**
```{r}

get_dist_params <- function(dist, data) { 
  
  mu <- mean(data); var <- var(data)
  # For the beta distribution, run Newton-Raphson 20 times 
  if (dist=="beta") { 
    x <- data; a2 <- 1; b2 <- 1
    for (i in 1:20){
    # Matrices to hold estimates
      estim <- matrix(c(a2,b2), nrow=2)
    # Compute the partials with respect to alpha and beta 
      f<- matrix(c(digamma(a2+b2) - digamma(a2) + mean(log(x)),
                   digamma(a2+b2)- digamma(b2)+ mean(log(1-x))), nrow=2) 
      
      # Compute partial derivatives by solving the Jacobian matrix
      J <- solve(matrix(c(trigamma(a2+b2)-trigamma(a2), 
                          trigamma(a2+b2), trigamma(a2+b2),
                          trigamma(a2+b2)-trigamma(b2)), nrow=2,ncol=2)) 
      
      f2 <- estim - J%*%f
      a2 <- f2[1,]
      b2 <- f2[2,]
    }
    cat("Beta dist alpha:",a2,'\n') 
    cat("Beta dist beta: ",b2,'\n')
  }
  else if (dist == "bernoulli") {
    cat("Bernoulli p:", mu,'\n')
  }
  else if (dist=="binomial") { 
    cat("Binomial p: ", mu/length(data),'\n')
  }
  else if (dist == "exponential") {
    cat("Exponential rate value:", 1/mu,'\n') 
  }
  # For gamma distribution, we run Newton-Raphson for 60 iterations
  else if (dist=="gamma") { 
    x <- data; n <- length(data); a <- 1
    
    for(i in 1:60){ 
      a <- a - f1(a,n,x)/f2(a,n) 
    }
    # Beta value
    b <- mean(x)/a
    cat("Gamma alpha: ",a,'\n') 
    cat("Gamma beta: ",1/b,'\n') 
  }
  else if (dist=="geometric") {
    p <- 1/mu
    cat("Geometric p: ", p,'\n')
  }
  else if (dist=="multinomial") {
    n_row <- nrow(data); prob <- c(0,0,0,0);  
    p <- data/length(data) 
    prob <- rowSums(p)
    cat('Multinomial prob: ',prob,'\n')
  }
  else if (dist=="multivariate_normal") {
    mu <- colMeans(data) 
    x_sub_mean <- data - mu
    cov <- matrix(c(0,0,0,0),2,2)
    for (i in 1:10000) {
      prod <- x_sub_mean[i,] %*% t(x_sub_mean[i,])
      cov <- cov + prod 
    }
    covariance <- round(cov / 10000, 3)
    cat(covariance,'\n')
    cat("Multivariate mean: ", mu," Multivariate Sigma: ", covariance,'\n')
  }
  else if (dist=="normal") {
    cat("Normal Mean: value: ", mu,'\n')
    cat("Normal Standard deviation: ", sqrt(var),'\n') 
  }
  else if (dist=="poisson") {
    cat("Poisson Lambda: ", mu,'\n')
  }
  else if (dist=="uniform") {
    cat("Uniform a: ",min(data)," Uniform b: ", max(data),'\n')
  }
}
```
**This function is used to calculate parameters for**
**different distributions of data using mean and variance**
```{r}
options(warn=-1)
get_kstest <- function(dist, data) { 
  
  mu <- mean(data); var <- var(data)
  d<-c(0)
  # For the beta distribution, run Newton-Raphson 20 times 
  if (dist=="beta") { 
    x <- data; a2 <- 1; b2 <- 1
    for (i in 1:20){
    # Matrices to hold estimates
      estim <- matrix(c(a2,b2), nrow=2)
    # Compute the partials with respect to alpha and beta 
      f<- matrix(c(digamma(a2+b2) - digamma(a2) + mean(log(x)),
                   digamma(a2+b2)- digamma(b2)+ mean(log(1-x))), nrow=2) 
      
      # Compute partial derivatives by solving the Jacobian matrix
      J <- solve(matrix(c(trigamma(a2+b2)-trigamma(a2), 
                          trigamma(a2+b2), trigamma(a2+b2),
                          trigamma(a2+b2)-trigamma(b2)), nrow=2,ncol=2)) 
      
      f2 <- estim - J%*%f
      a2 <- f2[1,]
      b2 <- f2[2,]
    }
    for(i in 1:100){
      yboot <- rbeta(10000, a2, b2)
      d[i] <- ks.test(yboot,"pbeta",a2,b2)$statistic  #test against beta
    }
    d0 = ks.test(data,"pbeta",2, 8)$statistic
    cat("d0 = ",d0,";d*mean = ", mean(d))
  }
  else if (dist == "bernoulli") {
    p <- mu
      for(i in 1:100){
        yboot <- rbinom(1000,1,p)
        d[i] <- ks.test(yboot,"pbinom",1,p)$statistic  #test against beta
      }
    d0 <- ks.test(data,"pbinom",1, 0.75)$statistic
    cat("d0 = ",d0,";d*mean = ", mean(d))
  }
  else if (dist=="binomial") { 
    p <- mu/(10*length(data))
      for(i in 1:100){
        yboot <- rbinom(100,1000,p)
        d[i] <- ks.test(yboot,"pbinom",1000,p)$statistic  #test against beta
      }
    d0 <- ks.test(data,"pbinom",1000, 0.75)$statistic
    cat("d0 = ",d0,";d*mean = ", mean(d))
  }
  else if (dist == "exponential") {
    a <- 1/mu
      for(i in 1:100){
        yboot <- rexp(100000,a)
        d[i] <- ks.test(yboot,"pexp",a)$statistic  #test against beta
      }
    d0 <- ks.test(data,"pexp",5)$statistic
    cat("d0 = ",d0,";d*mean = ", mean(d))
  }
  # For gamma distribution, we run Newton-Raphson for 60 iterations
  else if (dist=="gamma") { 
    x <- data; n <- length(data); a <- 1
    
    for(i in 1:60){ 
      a <- a - f1(a,n,x)/f2(a,n) 
    }
    # Beta value
    b <- mean(x)/a
    d0 = ks.test(data,"pgamma",2, 3)$statistic
    for(i in 1:100){
      yboot <- rgamma(100000, a, b)
      d[i] <- ks.test(yboot,"pgamma",a,b)$statistic
    }
    cat("d0 = ",d0,";d*mean = ", mean(d))
  }
  else if (dist=="geometric") {
    p <- 1/mu
    d0 = ks.test(data,"pgeom",0.25)$statistic
    for(i in 1:100){
      yboot <- rgeom(100000,p)
      d[i] <- ks.test(yboot,"pgeom",p)$statistic
    }
    cat("d0 = ",d0,";d*mean = ", mean(d))
  }
  else if (dist=="multinomial") {
    n_row <- nrow(data); prob <- c(0,0,0,0);  
    p <- data/length(data) 
    prob <- rowSums(p)
    cat('Multinomial prob: ',prob,'\n')
  }
  else if (dist=="multivariate_normal") {
    mu <- colMeans(data) 
    x_sub_mean <- data - mu
    cov <- matrix(c(0,0,0,0),2,2)
    for (i in 1:10000) {
      prod <- x_sub_mean[i,] %*% t(x_sub_mean[i,])
      cov <- cov + prod 
    }
    covariance <- round(cov / 10000, 3)
    cat(covariance,'\n')
    cat("Multivariate mean: ", mu," Multivariate Sigma: ", covariance,'\n')
  }
  else if (dist=="normal") {
    std <- sqrt(var)
    d0 = ks.test(data,"pnorm",20,2)$statistic
    for(i in 1:100){
      yboot <- rnorm(100000,mu, std)
      d[i] <- ks.test(yboot,"pnorm",mu,std)$statistic
    }
    cat("d0 = ",d0,";d*mean = ", mean(d))
  }
  else if (dist=="poisson") {
    d0 <- ks.test(data,"ppois",3)$statistic
    l <- mu
     for(i in 1:100){
       yboot <- rpois(40000,lambda = l)
       d[i] <- ks.test(yboot,"ppois",l)$statistic
     }
     cat("d0 = ",d0,";d*mean = ", mean(d))
  }
  else if (dist=="uniform") {
    d0 <- ks.test(data,"punif",1,100)$statistic
    a <- min(data)
    b <- max(data)
     for(i in 1:100){
       yboot <- runif(1000000,a,b)
       d[i] <- ks.test(yboot,"punif",a,b)$statistic
     }
     cat("d0 = ",d0,";d*mean = ", mean(d))
  }
}
```

**Bernoulli Distribution**
```{r}
options(warn=-1)
bernoulli_data <- rbinom(1000, 1, 0.75) 
dist <- 'bernoulli' 
p <- mean(bernoulli_data)
get_dist_params(dist,bernoulli_data)
get_kstest(dist,bernoulli_data)
```

**Beta Distribution**
```{r}
options(warn=-1)
beta_data <- rbeta(10000, 2, 8) 
dist <- 'beta'
get_dist_params(dist,beta_data)
get_kstest(dist,beta_data)

```


**Binomial Distribution**
```{r}
options(warn=-1)
binom_data <- rbinom(100, 1, 0.5) 
dist <- 'binomial' 
get_dist_params(dist,binom_data)
get_kstest(dist,binom_data)
#bayesian
palpha=1
pbeta=1
r=1
#posterior parameters
poalpha=palpha+sum(binom_data)
pobeta=pbeta+length(binom_data)-sum(binom_data)
psample=rbeta(1000,poalpha,pobeta)
plot(density(psample),main='Binomial')
```

**Exponential Distribution**
```{r}
options(warn=-1)
exp_data <- rexp(100, 1) 
dist <- 'exponential'
get_dist_params(dist,exp_data)
get_kstest(dist,exp_data)
#Bayesian
#prior distribution parameters
palpha=1
pbeta=1
#posterior parameters
poalpha=palpha+length(exp_data)
pobeta=pbeta+sum(exp_data)
psample=rgamma(1000,poalpha,pobeta)
plot(density(psample),main='Exponential')
```

**Gamma Distribution**
```{r}
options(warn=-1)
gamma_data <- rgamma(10000, 2, 3) 
dist <- 'gamma'
get_dist_params(dist,gamma_data)
get_kstest(dist,gamma_data)
```

**Geometric Distribution**
```{r}
options(warn=-1)
geom_data <- rgeom(100000, 0.25)
dist <- 'geometric'
get_dist_params(dist,geom_data)
get_kstest(dist,geom_data)
```


**Multinomial Distribution**
```{r}
p = c(0.2,0.3,0.5)
data <- rmultinom(1000,c(1,2,7),p) 
dist <- 'multinomial'
get_dist_params(dist, data)
#Bayesian
#prior distribution parameters
palpha=p
x=c(1,2,7)
#posterior parameters
poalpha=palpha+x
psample=rDirichlet(1000,poalpha)
plot(density(psample),main='Multinomial')
```

**Multivariate Normal Distribution**
```{r}
Sum <- matrix(c(9,6,6,16),2,2)
data <- mvrnorm(n = 10000, c(4, 5), Sum) 
dist <- 'multivariate_normal'
get_dist_params(dist, data)

```

**Normal Distribution**
```{r}
norm_data <- rnorm(1000, 0, 2) 
dist <- 'normal'
get_dist_params(dist,norm_data)
get_kstest(dist,norm_data)
#Bayesian
set.seed(12345)
norm<-runif(1000)
#prior distribution parameters
palpha=0.1
pbeta=0.1
mu=0.1
t=1/var(norm)
r=0.01
#posterior parameters
#conditional
pomu=(t*mu+length(norm)*mean(norm)*r)/(t+length(norm)*r)
pot=(t+length(norm))*r
psample1=rnorm(1000,pomu,1/sqrt(pot))
plot(density(psample1),main='Conditional_Normal')
#Marginal
poalpha=palpha+length(norm)/2
pobeta=pbeta+(var(norm)*length(norm))/2+(t*length(norm)*(mean(norm)-mu)**2)/(2*(t+length(norm)))
psample2=rgamma(1000,poalpha,pobeta)
plot(density(psample2),main='Marginal_Normal')
```

**Poisson Distribution**
```{r}
options(warn=-1)
poisson_data <- rpois(1000, 1) 
dist <- 'poisson' 
    
get_dist_params(dist,poisson_data)
get_kstest(dist,poisson_data)
#Bayesian
#prior distribution parameters
palpha=1
pbeta=1
#posterior parameters
poalpha=palpha+sum(poisson_data)
pobeta=pbeta+length(poisson_data)
psample=rgamma(1000,poalpha,pobeta)
plot(density(psample),main='Poisson')
```
**Uniform Distribution**
```{r}
options(warn=-1)
uniform_data <- runif(1000000, 1, 100) 
dist <- 'uniform' 
get_dist_params(dist,uniform_data)
get_kstest(dist,uniform_data)
```












