---
title: 'Assignment 4: Bootstrap and Jackknife'
author: 'GROUP PROJECT: SriKaavya Toodi(st938), Rithvik (ra830),Aravind reddy(ad1452),Sowmya(sa1607)'
date: "03/10/2020"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Build a bias, standard deviation, and  confidence interval estimator for an arbitrary statistic (named statfunc) based on the bootstrap (use 10000 =nboot) and the jackknife

# bootstrap function
```{r Bootstrap}
bootstrap <-function(vec0,statfunc=mean,nboot=10000,alpha=0.05)
{
  #Extracting sample size, mean and standard deviation from the original data
  n0<-length(vec0) 
  mean0<-mean(vec0)
  
  jack = Jackknife(vec0,statfunc) #Jackknife function is called over here with input vector Vec0
  sd0<-jack$jacksd #jackknife Standard Deviation
  bias0<-jack$jackbias #jackknife bias
  
  #create a vector to store the location of the bootstrap studentized deviation vector

  boot_pivot_vec<-NULL #vector for the bootpivotal values
  boot_bias_vec<-NULL #vector for the bootbias values 
  jacck_bias_vec<-NULL #vector for the jackbias values 
  jack_sd_vec<-NULL #vector for the jacksd values 
  boot_percentile_vec<-NULL #vector for the bootpercentile values 
  
  
  for(i in 1:nboot)
  { 
    vecb<-sample(vec0,replace=T) 
    meanb<-statfunc(vecb)
    sdb<-Jackknife(vecb,statfunc)$jacksd
    boot_pivot_vec<-c(boot_pivot_vec,(meanb-mean0)/(sdb/sqrt(n0))) 
    boot_bias_vec<-c(boot_bias_vec,meanb-mean0) 
    boot_percentile_vec = c(boot_percentile_vec,meanb)
  }
  
  # bias in the statistic
  bootbias<-mean(boot_bias_vec)
  
  
  #Calculating lower and upper quantile of the bootstrap distribution
  lq<-quantile(boot_pivot_vec,alpha/2)
  uq<-quantile(boot_pivot_vec,1-alpha/2)
  
  lqpercentile = quantile(boot_percentile_vec,alpha/2)
  uqpercentile = quantile(boot_percentile_vec, 1-alpha/2)
  
  sdboot = sd(boot_percentile_vec) 
  LB<-mean0-(sd0/sqrt(n0))*uq
  UB<-mean0-(sd0/sqrt(n0))*lq
  
  LBpercentile<-lqpercentile
  UBpercentile<-uqpercentile
  
  NLB<-mean0-(sd0/sqrt(n0))*qt(1-alpha/2,n0-1)
  NUB<-mean0+(sd0/sqrt(n0))*qt(1-alpha/2,n0-1)
  
  BNCLB = mean0 -sdboot*qt(1-alpha/2,n0-1)
  BNCUB = mean0 +sdboot*qt(1-alpha/2,n0-1)
  
  list(bootstrap.pivotal.confidence.interval=c(LB,UB),normal.confidence.interval.jackknife=
c(NLB,NUB),

bootstrap.percentile.confidence.interval=c(LBpercentile,UBpercentile),boot.normal.confidence.interval = c(BNCLB,BNCUB),bootbias=bootbias,jackbias = bias0) 
}
```

# Jackknife Function
```{r Jackknife}
Jackknife<-function(v1,statfunc=sd){
  
  n1<-length(v1) 
  jackvec<-NULL
  mu0<-statfunc(v1) 
  
  for(i in 1:n1)
  { 
    mua<-statfunc(v1[-i]) 
    jackvec<-c(jackvec, n1*(mu0)-(n1-1)*mua) 
  } 
  
  jackbias<-mean(jackvec)-mu0
  jacksd<-sd(jackvec) 
  
  list(mu0=mu0,jackbias=jackbias,jacksd=jacksd)
}

```

# Simulation Function
```{r SimFunc}
Sim.func<-function(mu.val=3,n=30,nsim=1000)
{
 #empty vectors for the bootstrap, normal and Jackknife confidence interval
 cvec.boot_pivotal<-NULL #coverage vector for boot pivotal confidence interval
 cvec.norm_jackknife<-NULL #coverage vector for normal confidence interval(jackknife)
 cvec.boot_normal<-NULL #coverage vector for boot normal confidence interval
 cvec.boot_percentile = NULL #coverage vector for boot percentile confidence interval

 #Calculating the mean of the lognormal distribution
 mulnorm<-(exp(mu.val+1/2))
 
 #run simulation
 for(i in 1:nsim){ #i is current simulation number

 if((i/10)==floor(i/100)){ print(i)
 #let me know computer hasn’t died
 }

 #sample the simulation vector
 vec.sample<-rlnorm(n,mu.val)

 #bootstrap it
 boot.trial<-bootstrap(vec.sample) #Call for the bootstrap function to obtain required confidence intervals

 boot_pivotal.conf<-boot.trial$bootstrap.pivotal.confidence.interval
 norm_jackknife.conf<-boot.trial$normal.confidence.interval.jackknife
 boot_percentile.conf<-boot.trial$bootstrap.percentile.confidence.interval
 boot_normal.conf<-boot.trial$boot.normal.confidence.interval

 #calculate if confidence intervals include mu

 #count up the coverage by the bootstrap pivotal interval
 cvec.boot_pivotal<-
c(cvec.boot_pivotal,(boot_pivotal.conf[1]<mulnorm)*(boot_pivotal.conf[2]>mulnorm))

 #count up the coverage by the normal confidence interval(jackknife)
 cvec.norm_jackknife <- c(cvec.norm_jackknife,(norm_jackknife.conf[1]<mulnorm)*(norm_jackknife.conf[2]>mulnorm))

 #count up the coverage by the bootstrap percentile interval
 cvec.boot_percentile<- c(cvec.boot_percentile,(boot_percentile.conf[1]<mulnorm)*(boot_percentile.conf[2]>mulnorm))

 #count up the coverage by the bootstrap normal interval
 cvec.boot_normal<-
c(cvec.boot_normal,(boot_normal.conf[1]<mulnorm)*(boot_normal.conf[2]>mulnorm))
 }
 #calculate and output coverage probability estimates
list(boot_pivotal.coverage=(sum(cvec.boot_pivotal)/nsim),norm_jackknife.coverage=(sum(cvec.norm_jackknife)/nsim),

boot_percentile.coverage=(sum(cvec.boot_percentile)/nsim),boot_normal.coverage=(sum(cvec.boot_normal)/nsim)) 
}
```

The above function calls the bootstrap and Jackknife function nsim times.

The lower and upper bound for the 4 confidence intervals are obtained. 


```{r Sim1}
#Sim for n=10
Sim.func(3,10,1000)
```

```{r Sim2}
#Sim for n=30
Sim.func(3,30,1000)
```

```{r Sim3}
#Sim for n=100
Sim.func(3,100,1000)
```




## Estimate the bias by computing the difference between the standard deviation of the sample data and the standard deviation of the original data.

# bootstrap:

```{r BootBiasEst}
BootstrapBias<-function(vec0,nsim=1000)
{
  n0<-length(vec0) 
  mean0<-mean(vec0)
  sd0<-sqrt(var(vec0)) 
  
  bootvec<-NULL
  bootbiasvec<-NULL 
  
  for(i in 1:n0){
    vecb<-sample(vec0,replace=T) 
    nb<-length(vecb)
    meanb<-mean(vecb)
    sdb<-sqrt(var(vecb))
  }
	
  final_bias_boot<-(sdb/nb)-(sd0/n0) #nsim is no.of simulations

	list(bootstrap_bias = final_bias_boot)
}
```


### Next, for Jackknife:
```{r JackBiasEst}
JackknifeBias<-function(v1,statfunc=sd,nsim=1000)
{
  n1<-length(v1) 
  jack_vec<-NULL 
  mu0<-statfunc(v1) 
  for(i in 1:n1){
    mua<-statfunc(v1[-i])
    jack_vec<-c(jack_vec, n1*(mu0)-(n1-1)*mua) 
  }
  jackbias <- mean(jack_vec)-mu0
  jacksd<-sd(jack_vec)

  final_bias_jack<-(jacksd/n1)-(mu0/n1) #nsim is no.of simulations
  list(jackknife_bias = final_bias_jack)
}
```



# Calling the functions with sample data
```{r BootVsJackBiasComp}
Sim.bias<-function(mu.val=3,n=30,nsim=1000)
{
  biasBootstrap<-NULL
  biasJackknife<-NULL
  
  #run simulation
  for(i in 1:nsim)
  {
    vec.sample<-rlnorm(n,mu.val) 
    
    boot.list<-BootstrapBias(vec.sample) 
    jack_knife.list<-JackknifeBias(vec.sample) 
    
    biasBootstrap<-c(biasBootstrap,boot.list$bootstrap_bias)
    biasJackknife<-c(biasJackknife,jack_knife.list$jackknife_bias)
  }
  
  list(biasBoot = biasBootstrap, biasJack = biasJackknife)
}
```

# Running the Simulations function
```{r BiasCompSim}
var<-Sim.bias(3,30,1000)
var
```


```{r BootPlot}
# Line Plot

d <- var$biasBoot
dj <-var$biasJack

plot(d,main ="Comparision between the bootstrap and the jacknife bias",type="l",col="red")
lines(dj,col="blue")
```



