---
title: "Assignment-6"
author: "MLE Estimators: SriKaavya(st938), Rithvik (ra830), Aravind (ad1452), Sowmya(sa1607)"
date: "10-21-2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Function for Maximum likely hood estimation.

Method to estimate parameters for different distributions is below. The input parameters are:- 
"distribution_type" : Type of distribution. To be given as a string input
"data_input" : Data generated using the specific distribution to be given as input

Mean and variance and given data is calculated and using maximum likelihood estimation, calculations will be done to estimate specific parameter values. These will be printed after calculations.

```{r parameter func, echo=TRUE}
library(purrr)
library(MASS)
 Parameter_finder <- function( distribution_type , data_input ) {
  # This function is used to calculate parameters for
  #different distributions of data using mean and variance
  mean_data = mean(data_input) 
  var_data = var(data_input)
  
  if ( distribution_type == "Poisson") {
    print(paste0("Poisson distribution parameters - Lambda value is ", mean_data))
  } 
  
  else if (  distribution_type == "Bernoulli") {
    print(paste0("Bernoulli distribution parameters - p value is", mean_data))
  }
  else if (  distribution_type == "Exponential") {
    print(paste0("Exponential distribution parameters - rate value is ", 1/mean_data))
  }
   else if (distribution_type == "Geometric") {
    p = 1/mean_data
    print(paste0("Geometric distribution parameters - p value is ", p))
   }
  else if (  distribution_type == "Normal") {
    print(paste0("Normal distribution parameters - Mean value is ", mean_data))
    print(paste0("Standard deviation is ", sqrt(var_data)))
  }
  else if (  distribution_type == "Binomial") {
    print(paste0("Binomial distribution parameters - p value is ",
                 mean_data/length(data_input)))
  }
  else if (  distribution_type == "Uniform") {
    print(paste0("Uniform distribution parameters - a value is ", min(data_input),
                 " b value is ", max(data_input)))
  }
  
   else if (  distribution_type == "Gamma") {
    x<- data_input
    n = length(data_input)
    a <- 1
   # Defining derivative functions for newton-raphson
   f1 <- function(a) -n*digamma(a)- n*log(mean(x)) + n*log(a) + sum(log(x))
   f2 <- function(a) -n*trigamma(a) + n/a
  
   #Newton-Raphson for 60 iterations
   for( i in 1:60) a <- a - f1(a)/f2(a)
   # Beta value
   b <- mean(x)/a
  
   print(paste0("Estimate of alpha is ",a))
   # As input to R is inverse of beta
   print(paste0("Estimate of beta is ",1/b))
   }
  
  
  else if (  distribution_type == "Beta") {
    x <- data_input
   # Initial guesses
   a2 <- 1
   b2<- 1
   # Running Newton-Raphson 20 times
   for (i in 1:20){
     # Matrices to hold estimates
   parm <- matrix(c(a2,b2), nrow=2)
   # Two components by partial derivatives with respect to alpha and beta
   f<- matrix(c(digamma(a2+b2) - digamma(a2) + mean(log(x)),
                digamma(a2+b2)- digamma(b2)+ mean(log(1-x))), nrow=2)
   # Solving these using jacobian matric
   J <- solve(matrix(c(trigamma(a2+b2)-trigamma(a2),
                       trigamma(a2+b2), trigamma(a2+b2),
                       trigamma(a2+b2)-trigamma(b2)), nrow=2,ncol=2))
   # Matrix multiplication
   f2 <- parm -  J%*%f
   # Getting estimates and putting them back inside
   a2 <- f2[1,] 
   b2 <- f2[2,]
   }
   
   print(paste0("Estimate of alpha is ",a2))
   # As input to R is inverse of beta
   print(paste0("Estimate of beta is ",b2))
  
  }
  
  else if (distribution_type=="multinomial") 
  {
    n_row = nrow(data_input)
    prob = c(0,0,0,0)
    p=data_input/length(data_input)
    for(i in 1:n_row)
      prob[i] = sum(p[i,])
    print(prob)
    #print(paste0("Uniform multinomial parameters - n ", list (n=n)))
    #print(paste0("Uniform multinomial parameters - p ",prob))
  }
  else if (distribution_type=="multivariatenormal") 
  {
    mean_data <- colMeans(data_input)
    x_sub_mean <- data-mean_data
    cov <- matrix(c(0,0,0,0),2,2)
    for (i in 1:10000) {
      prod <- x_sub_mean[i,] %*% t(x_sub_mean[i,])
      cov <- cov + prod
    }
    covariance <- round(cov / 10000, 2)
    print(covariance)
    print(paste0("Uniform distribution parameters - mean value is ", mean_data,
                 " Sigma value is ", covariance))
  }

 }
  
```

## Testing distributions

Testing different distributions below:

```{r Testing, echo=TRUE}
# Bernoulli is Binomial with size as 1
 bernoulli_data <- rbinom(1000, 1, 0.75)
 distribution_type = 'Bernoulli'
 Parameter_finder(distribution_type,bernoulli_data)
 
 #Binomial distribution
 binom_data <- rbinom(100, 1000, 0.75)
 distribution_type = 'Binomial'
 Parameter_finder(distribution_type,binom_data)
 
 # Geometric distribution. Here parameter couln't 
 # be estimated even with really high value of N
 geom_data <- rgeom(100000, 0.25)
 distribution_type = 'Geometric'
 Parameter_finder(distribution_type,geom_data)
 
 
 # Poisson distribution. Lamdba is accurately estimated
 poisson_data <- rpois(40000, lambda = 3)
 distribution_type = 'Poisson'
 Parameter_finder(distribution_type,poisson_data)
 
 
 #Uniform distribution data.
 #Here a and B are accurately estimated if gap between them is large enough.
 #If not,it seems inaccurate
  uniform_data <- runif(1000000, 1, 100)
 distribution_type = 'Uniform'
 Parameter_finder(distribution_type,uniform_data)
 
 # Normal distribution
 norm_data <- rnorm(100000, 20, 2)
 distribution_type = 'Normal'
 Parameter_finder(distribution_type,norm_data)
 
 # Exponential distribution
 exp_data <- rexp(100000, 5)
 distribution_type = 'Exponential'
 Parameter_finder(distribution_type,exp_data)
 
 # Gamma distribution
 gamma_data <- rgamma(10000, 2, 3)
 distribution_type = 'Gamma'
 Parameter_finder(distribution_type,gamma_data)

 # Beta distribution
 beta_data <- rbeta(10000, 2, 8)
 distribution_type = 'Beta'
 Parameter_finder(distribution_type,beta_data)

 #Multinomial Distribution
  p = c(0.20,0.40,0.05,0.30)
  data = rmultinom(10000,size=4,p)
  Parameter_finder("multinomial", data)
 
 # Multi variate normal distribution
 Sum = matrix(c(9,6,6,16),2,2)
 data = mvrnorm(n = 10000, c(4, 5), Sum)
 distribution_type = 'multivariatenormal'
 Parameter_finder(distribution_type, data)
 
```

