---
title: "Assignment_5"
author: "GROUP PROJECT: SriKaavya Toodi(st938), Rithvik (ra830),Aravind reddy(ad1452),Sowmya(sa1607)"
date: "10/13/2020"
output: 
  pdf_document: 
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Function to estimate

Method to estimate parameters for different distributions is below. Input parameters are 
"distribution_type" : Type of distribution. To be given as a string input
"data_input" : Data generated using the specific distribution to be given as input

Mean and variance and given data is calculated and using Method of moments, calculations will be done to estimate specific parameter values. These will be printed after calculations. 

We've to note that calculating variance for the data set implicitly has second moment in it. Using variance for estimating a parameter is almost similar to using second moment of the given data.


```{r parameter func, echo=TRUE}
library(purrr)
library(MASS)
 Parameter_finder <- function( distribution_type , data_input , k) {
  # This function is used to calculate parameters for
  #different distributions of data using mean and variance
  if(!distribution_type=="multinomial"){
   print(head(data_input))
  }
  mean_data = mean(data_input) 
  var_data = var(data_input)
  
  if ( distribution_type == "Poisson") {
    print(paste0("Poisson distribution parameters - Lambda value is ", mean_data))
  } 
  else if (  distribution_type == "Point mass at a") {
    print(paste0("Point mass at a distribution parameters - a value is", mean_data))
  } 
  else if (  distribution_type == "Bernoulli") {
    print(paste0("Bernoulli distribution parameters - p value is", mean_data))
  } 
  else if (  distribution_type == "Binomial") {
    p = var_data/mean_data
    p = 1-p
    n = mean_data/p
    print(paste0("Binomial distribution parameters - p value is ",p))
    print(paste0("N value is ", n))
  }
  else if (  distribution_type == "Geometric") {
    p = 1/mean_data
    print(paste0("Geometric distribution parameters - p value is ", p))
  }
  else if (  distribution_type == "Uniform") {
    #sub_ba = (var_data/mean_data) * 6
    #sum_ab = mean_data * 2
    b = mean_data + (sqrt(3*var_data))
    a = mean_data - (sqrt(3*var_data))
    print(paste0("Uniform distribution parameters - a value is ", a,", b value is ", b))
  }
  else if (  distribution_type == "Normal") {
    print(paste0("Normal distribution parameters - Mean value is ", mean_data))
    print(paste0("Standard deviation is ", sqrt(var_data)))
  }
  else if (  distribution_type == "Exponential") {
    print(paste0("Exponential distribution parameters - rate value is ", 1/mean_data))
  }
  else if (  distribution_type == "Gamma") {
    beta = var_data/mean_data
    alpha = mean_data/beta
    print(paste0("Gamma distribution parameters - alpha value is", alpha))
    print(paste0("Beta value is ",beta, " and rate is ", 1/beta))
  }
  else if (  distribution_type == "Beta") {
    m <- mean_data
    v <- var_data
    alpha <- m*((m*(1-m)/v)-1)
    beta <- alpha*(1-m)/m
    print(paste0("Beta distribution parameters - alpha value is", alpha))
    print(paste0("Beta value is ",beta))
  }
  else if (  distribution_type == "tdist") {
    v = 2*var_data/(var_data-1)
    print(paste0("t distribution distribution parameters -v is ", v))
  }
  else if (  distribution_type == "Chi Square") {
    print(paste0("Chi Square distribution parameters - p value is ", mean_data))
  }

  else if (distribution_type=="multinomial") 
  {
    n_row = nrow(data_input)
    prob = c(0,0,0,0)
    for(i in 1:n_row)
      p[i]<-1-((var(data_input[i,]))/mean(data_input[i,]))
    n = sum(rowMeans(data_input))/sum(p[1:n_row])
    print(list(n=n, p=p))
  }
  else if (distribution_type=="multivariatenormal") 
  {
    mean = colMeans(data_input)
    Sigma = var(data_input)
    print(list(mu_val=mean, summation=Sigma))
  }else if (distribution_type=="hypergeometric") {
    N <- length(data_input)
    p <- mean_data/k
    product <- var_data/(k * p * (1-p))
    m_plus_n <- (k-product)/(1-product)
    m <- p * m_plus_n
    print(paste0("N value is ", N, " Subclass 1(m) value is ", m, " Subclass 2(n) value is ", m_plus_n-m))
  }
  #print(paste0('Mean of data is ', mean_data))
  #print(paste0('Variation of data is ', var_data))
 }
```

## Testing distributions

Testing different distributions below

```{r Testing, echo=TRUE}
# Bernoulli is Binomial with size as 1
 bernoulli_data <- rbinom(1000, 1, 0.75)
 distribution_type = 'Bernoulli'
 Parameter_finder(distribution_type,bernoulli_data)
 
 #Binomial distribution
 binom_data <- rbinom(100, 1000, 0.75)
 distribution_type = 'Binomial'
 Parameter_finder(distribution_type,binom_data)
 
 # Geometric distribution. Here parameter couln't be estimated even with really high value of N
 geom_data <- rgeom(100000, 0.25)
 distribution_type = 'Geometric'
 Parameter_finder(distribution_type,geom_data)
 
 
 # Poisson distribution. Lamdba is accurately estimated
 poisson_data <- rpois(40000, lambda = 3)
 distribution_type = 'Poisson'
 Parameter_finder(distribution_type,poisson_data)
 
 
 #Uniform distribution data. Here a and B are accurately estimated if gap between them is large enough.
 #If not,it seems inaccurate
  uniform_data <- runif(1000000, 1, 100)
 distribution_type = 'Uniform'
 Parameter_finder(distribution_type,uniform_data)
 
 # Normal distribution
 norm_data <- rnorm(100000, 20, 2)
 distribution_type = 'Normal'
 Parameter_finder(distribution_type,norm_data)
 
 # Exponential distribution
 exp_data <- rexp(100000, 5)
 distribution_type = 'Exponential'
 Parameter_finder(distribution_type,exp_data)
 
 # Gamma distribution
 gamma_data <- rgamma(10000, 2, 3)
 distribution_type = 'Gamma'
 Parameter_finder(distribution_type,gamma_data)
 
 # Beta distribution
 beta_data <- rbeta(10000, 2, 8)
 distribution_type = 'Beta'
 Parameter_finder(distribution_type,beta_data)
 
 # Student T distribution
 tdist_data = rt(100000, 6)
 distribution_type = 'tdist'
 Parameter_finder(distribution_type,tdist_data)
 
 # CHI-SQUARE dustribution
 chi2_data <- rchisq(10000, 5)
 distribution_type = 'Chi Square'
 Parameter_finder(distribution_type,chi2_data)
 
 #Multinomial Distribution
  p = c(0.20,0.40,0.05,0.30)
  data = rmultinom(1000,size=4,p)
  Parameter_finder("multinomial", data)
 
 # Multi variate normal distribution
 Sum = matrix(c(7,3,3,7),2,2)
 data = mvrnorm(n = 100000, rep(0, 2), Sum)
 distribution_type = 'multivariatenormal'
 Parameter_finder(distribution_type, data)
 
 # Hyper Geometric Distribution
 k <- 3
 data <- rhyper(40000, 10, 15, 3)
 distribution_type = 'hypergeometric'
 Parameter_finder(distribution_type, data, k)
 
```

