---
title: "Probability and Inference HW 3b"
author: "Yaniv Bronshtein(yb262), Vipul Ishwarchandra Gharde(vig4), Animesh Sharma(as3592),Wanying Mo(wm318), Fan Shen FS470,   "
date: "11/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Import the necessary libraries**
```{r}
library(purrr) 
library(MASS)
```

**Functions to calculate first and second derivatives for Newton Raphson method**
```{r}
f1 <- function(a,n,x){
  return(-n*digamma(a)- n*log(mean(x)) + n*log(a) + sum(log(x)))
}
f2 <- function(a,n){
  return(-n*trigamma(a) + n/a)
}
```

**This function is used to calculate parameters for**
**different distributions of data using mean and variance**
```{r}

get_dist_params <- function(dist, data) { 
  
  mu <- mean(data); var <- var(data)
  # For the beta distribution, run Newton-Raphson 20 times 
  if (dist=="beta") { 
    x <- data; a2 <- 1; b2 <- 1
    for (i in 1:20){
    # Matrices to hold estimates
      estim <- matrix(c(a2,b2), nrow=2)
    # Compute the partials with respect to alpha and beta 
      f<- matrix(c(digamma(a2+b2) - digamma(a2) + mean(log(x)),
                   digamma(a2+b2)- digamma(b2)+ mean(log(1-x))), nrow=2) 
      
      # Compute partial derivatives by solving the Jacobian matrix
      J <- solve(matrix(c(trigamma(a2+b2)-trigamma(a2), 
                          trigamma(a2+b2), trigamma(a2+b2),
                          trigamma(a2+b2)-trigamma(b2)), nrow=2,ncol=2)) 
      
      f2 <- estim - J%*%f
      a2 <- f2[1,]
      b2 <- f2[2,]
    }
    cat("Beta dist alpha:",a2,'\n') 
    cat("Beta dist beta: ",b2,'\n')
  }
  else if (dist == "bernoulli") {
    cat("Bernoulli p:", mu,'\n')
  }
  else if (dist=="binomial") { 
    cat("Binomial p: ", mu/length(data),'\n')
  }
  else if (dist == "exponential") {
    cat("Exponential rate value:", 1/mu,'\n') 
  }
  # For gamma distribution, we run Newton-Raphson for 60 iterations
  else if (dist=="gamma") { 
    x <- data; n <- length(data); a <- 1
    
    for(i in 1:60){ 
      a <- a - f1(a,n,x)/f2(a,n) 
    }
    # Beta value
    b <- mean(x)/a
    cat("Gamma alpha: ",a,'\n') 
    cat("Gamma beta: ",1/b,'\n') 
  }
  else if (dist=="geometric") {
    p <- 1/mu
    cat("Geometric p: ", p,'\n')
  }
  else if (dist=="multinomial") {
    n_row <- nrow(data); prob <- c(0,0,0,0);  
    p <- data/length(data) 
    prob <- rowSums(p)
    cat('Multinomial prob: ',prob,'\n')
  }
  else if (dist=="multivariate_normal") {
    mu <- colMeans(data) 
    x_sub_mean <- data - mu
    cov <- matrix(c(0,0,0,0),2,2)
    for (i in 1:10000) {
      prod <- x_sub_mean[i,] %*% t(x_sub_mean[i,])
      cov <- cov + prod 
    }
    covariance <- round(cov / 10000, 3)
    cat(covariance,'\n')
    cat("Multivariate mean: ", mu," Multivariate Sigma: ", covariance,'\n')
  }
  else if (dist=="normal") {
    cat("Normal Mean: value: ", mu,'\n')
    cat("Normal Standard deviation: ", sqrt(var),'\n') 
  }
  else if (dist=="poisson") {
    cat("Poisson Lambda: ", mu,'\n')
  }
  else if (dist=="uniform") {
    cat("Uniform a: ",min(data)," Uniform b: ", max(data),'\n')
  }
}
```

# Let's test it now
**Bernoulli with size 1**
```{r}
# Bernoulli is Binomial with size as 1
bernoulli_data <- rbinom(1000, 1, 0.75) 
dist <- 'Bernoulli' 
get_dist_params(dist,bernoulli_data)
```
**Binomial Distribution**
```{r}
#Binomial distribution
binom_data <- rbinom(100, 1000, 0.75) 
dist <- 'Binomial' 
get_dist_params(dist,binom_data)
```
**Geometric Distribution**
```{r}
geom_data <- rgeom(100000, 0.25) 
dist <- 'geometric' 
get_dist_params(dist,geom_data)
```


**Poisson Distribution**
```{r}
poisson_data <- rpois(40000, lambda = 3) 
dist <- 'poisson' 
get_dist_params(dist,poisson_data)

```
**Uniform Distribution**
```{r}
uniform_data <- runif(1000000, 1, 100) 
dist <- 'uniform' 
get_dist_params(dist,uniform_data)
```

**Normal Distribution**
```{r}
norm_data <- rnorm(100000, 20, 2) 
dist <- 'normal'
get_dist_params(dist,norm_data)
```

**Exponential Distribution**
```{r}
exp_data <- rexp(100000, 5) 
dist <- 'exponential'
get_dist_params(dist,exp_data)
```
```{r}
gamma_data <- rgamma(10000, 2, 3) 
dist <- 'gamma' 
get_dist_params(dist,gamma_data)

```

**Beta Distribution**
```{r}
beta_data <- rbeta(10000, 2, 8) 
dist <- 'beta' 
get_dist_params(dist,beta_data)

```

**Multinomial Distribution**
```{r}
p = c(0.20,0.40,0.05,0.30)
data <- rmultinom(10000,size=4,p) 
dist <- 'multinomial'
get_dist_params(dist, data)
```


**Multivariate Normal Distribution**
```{r}
# Multi variate normal distribution
Sum <- matrix(c(9,6,6,16),2,2)
data <- mvrnorm(n = 10000, c(4, 5), Sum) 
dist <- 'multivariate_normal'
get_dist_params(dist, data)

```

