---
title: "Hurricane Analysis"
author: "Yaniv Bronshtein and Venkata Raju Datla"
date: "9/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Import the necessary libraries**
```{r}
library(tidyverse)
library(gridExtra) #For arranging ggplots
```

**Read in the hurricane data manually including both Category 4 and Category 5**
**Hurricane Data**
```{r}
path = "C:/Users/rajuk/OneDrive/Desktop/Class work/02 Fall 2021/03 Prob and stats/01 Home work/01 Hurricane modeling/category.csv"

hurricane_df <- read_csv("hurricane_data.csv")
```


**Create a calculated column for the decade to get per-decade data**
```{r}
hurricane_df['decade']=as.numeric(substr(hurricane_df$Year,1,3))*10
```

**Split the hurricane_df into category 4 and category 5 data. **
**In addition, generate the by decade median windspeed dataframes**
```{r}
cat4_df <- hurricane_df %>%
  filter(Category == 4)

cat5_df <- hurricane_df %>%
  filter(Category == 5)

cat_4df_by_decade <- cat4_df %>%
  group_by(decade) %>%
  summarise(Median_Wind = median(Wind))

cat_5df_by_decade <- cat5_df %>%
  group_by(decade) %>%
  summarise(Median_Wind = median(Wind))
```


**Create a function to generate scatter plots by year or by decade**
```{r}
get_windspeed_scatter <- function(data, title, by_year) {
  if(by_year) {
    return(
      ggplot(data=data, mapping = aes(x = Year, y = Wind)) +
        geom_point() +
        ggtitle(title) + 
        xlab("Year") +
        ylab("WindSpeed (km/h)") #+
    )
    
  } 
  else {
    return(ggplot(data=data, mapping = aes(x = decade, y = Median_Wind)) +
             geom_point() +
             ggtitle(title) +
             xlab("Decade") +
             ylab("Median Wind Speed (km/h)") #+
    )
    
  }
}
```


**Let's see some summary statistics first for the data**
```{r}
summary(cat4_df)
```

```{r}
summary(cat5_df)
```

```{r}
summary(cat_4df_by_decade)
```

```{r}
summary(cat_5df_by_decade)

```

**Generate the scatter plots for by year data**
```{r}
cat_4scatter <- get_windspeed_scatter(cat4_df, "Category 4 Hurricane Windspeed by Year", TRUE)

cat_5scatter <- get_windspeed_scatter(cat5_df, "Category 5 Hurricane Windspeed by Year", TRUE)

grid.arrange(cat_4scatter,cat_5scatter)
```


**Generate the scatter plots for by decade data**
```{r}
cat_4scatter_by_decade <- get_windspeed_scatter(cat_4df_by_decade, "Category 4 Hurricane Median Windspeed by Decade", FALSE)

cat_5scatter_by_decade <- get_windspeed_scatter(cat_5df_by_decade, "Category 5 Hurricane Median Windspeed by Decade", FALSE)

grid.arrange(cat_4scatter_by_decade,cat_5scatter_by_decade)
```

**Create a dataframe to get hurricane frequency data**
```{r}
dec_freq = hurricane_df %>% count(decade) %>% rename(freq="n")
dec_freq

```

**Generate the qqplot for the Poisson Distribution**
```{r}
ggplot(data = dec_freq,
       mapping = aes(sample = freq)) + 
  geom_qq(distribution = stats::qpois,
          dparams = list(lambda = mean(dec_freq$freq))) +
  geom_qq_line(distribution = stats::qpois,
               dparams = list(lambda = mean(dec_freq$freq)))

```

**Generate the qqplot for the Normal Distribution**
```{r}
ggplot(data = dec_freq,
       mapping = aes(sample = freq)) + 
  geom_qq() +
  geom_qq_line()

```

*Based on the plots Poisson is better because the data points are close to the 45 degree line*
